---
title: "For_Yang"
author: "Ruijuan Li"
date: "12/16/2018"
output: 
  html_document: 
    keep_md: yes
---

Yang has user study inforamtion from 18 people, including time and accurary value collected from three conditions. These 3 conditions are repeated measures, so I will use repeated measure to check whether there is significant difference among the three conditions interms of time and accuary. Benferroni correction will be used afterwards to correct for the P-value. In the end, need to make plot to show the result. 

```{r}
# load package 
library(tidyverse)
library(nlme)
library(reshape2)
library(car)

# read data 
user <- read.csv("~/Desktop/everyday_stats_bioinformatics/12-16-2018_repeated_measure_Anova/user.csv")
user %>% dim()
user %>% str()

# format data 
user_time <- 
user %>% 
  dplyr::select(name:t2) 

user_time <- user_time[1:18,]
user_time %>% head()

user_accuary <- 
user %>% 
  dplyr::select(name, a0:a2) 

user_accuary <- user_accuary[1:18,]
user_accuary %>% head()
```

### time 
```{r}
# make a linear model with only the intercept as the predictor for within-participant observations 
model1 <- lm(cbind(user_time$t0, user_time$t1, user_time$t2) ~ 1)
model1

conditions <- factor(colnames(user_time[2:4]), ordered = F) # factors for repeated measures variable 
conditions

model2=Anova(model1,idata=data.frame(conditions),idesign=~conditions,type="III") # use the repeated-measures factor as the "internal" part of the deign using ANOVA
# idata parameter is for the repeated-measures part of the data, the ‘idesign’ is where you specify the repeated part of the design. This is a one-way within design, so we have only the ‘Trials’ variable.

summary(model2,multivariate=F) # p-value is 1.654e-09, significant 

# pairwise 
c1=t.test(user_time$t0,user_time$t1,alternative="two.sided",mu=0,paired=T) 
c2=t.test(user_time$t0,user_time$t2,alternative="two.sided",mu=0,paired=T) 
c3=t.test(user_time$t1,user_time$t2,alternative="two.sided",mu=0,paired=T) 
c1; c2; c3

# adjust the p-value using Benferroni correction 
p.adjust(c(c1$p.value, c2$p.value, c3$p.value), method = "bonferroni", n = 18)   
```

### accuracy
```{r}
model3 <- lm(cbind(user_accuary$a0, user_accuary$a1, user_accuary$a2) ~ 1)
model3

conditions2 <- factor(colnames(user_accuary[2:4]), ordered = F) # factors for repeated measures variable 
conditions2

model4=Anova(model3,idata=data.frame(conditions2),idesign=~conditions2,type="III") # use the repeated-measures factor as the "internal" part of the deign using ANOVA

summary(model4,multivariate=F) # p-value is 1.654e-09, significant 

# pairwise 
c1.2=t.test(user_accuary$a0,user_accuary$a1,alternative="two.sided",mu=0,paired=T) 
c2.2=t.test(user_accuary$a0,user_accuary$a2,alternative="two.sided",mu=0,paired=T) 
c3.2=t.test(user_accuary$a1,user_accuary$a2,alternative="two.sided",mu=0,paired=T) 
c1.2; c2.2; c3.2

# adjust the p-value using Benferroni correction 
p.adjust(c(c1.2$p.value, c2.2$p.value, c3.2$p.value), method = "bonferroni", n = 18)   
```

### plot the data, I assume there should be a better way to get this (95% CI)
```{r}
mean <- 
  sapply(colnames(user)[2:7], function(i){
  round(mean(user[1:18,i]), digits = 2) 
}) 

stdr <- 
  sapply(colnames(user)[2:7], function(i){
  round(sqrt(var(user[1:18,i])/length(user[,i])), digits = 2) 
}) 

data <- 
as.data.frame(mean) %>% 
  cbind(as.data.frame(stdr)) %>% 
  mutate(max = mean + stdr, min = mean - stdr) 

data$condition <- colnames(user)[2:7]

data[1:3,] %>% 
  ggplot() + 
  geom_point(mapping=aes(x=condition,y=mean),stat="identity", size = 3, color = "darkorange") + 
  geom_errorbar(mapping=aes(x=condition,ymin=min,ymax=max), width = .1, position=position_dodge(0.01), color = "darkorange") + 
  ylim(0, 400)

data[4:6,] %>% 
  ggplot() + 
  geom_point(mapping=aes(x=condition,y=mean),stat="identity", size = 3, color = "blue") + 
  geom_errorbar(mapping=aes(x=condition,ymin=min,ymax=max), width = .1, position=position_dodge(0.01),  color = "blue") + 
  ylim(0, 1) 
```

