---
title: "08-05-2017"
author: "Ruijuan Li"
date: "8/5/2017"
output: 
  html_document: 
    keep_md: yes
---

### sum of square 

sum of square (R square explans how the model fit, low values of R squre indicate that the model is not adequate, however, high values of R square don't mean that the model is adequate.)

### load previous data 
```{r}
library(gee)
library(multcomp)

cholesterol = read.table("http://faculty.washington.edu/rhubb/sisg/SISG-Data-cholesterol.txt", header=T)
attach(cholesterol) # make variables global available in R studio 
dim(cholesterol) # 400 

# fit linear regression models for the association between triglycerdies and BMI
fit1 = lm(TG ~ BMI)
confint(fit1) # confidence interval 

# compute the predicted value and its 95% confidence interval for the mean value of TG at BMI=23 as well
# as for a new individual w/ BMI=23. 
predict(fit1, newdata = data.frame(BMI = 23), interval = "confidence") # predicted value and its 95% confidence interval for the mean value of TG at BMI=23
predict(fit1, newdata = data.frame(BMI = 23), interval = "prediction") # predicted value and its 95% confidence interval for a new individual w/ BMI=23
```

### simple linear regression assumption:
linearity, indepdence, normality, equal variance of residual 
```{r}
# plot of residuals VS. fitted values curvatures, heteroscendasticity? 
plot(fit1$fitted.values, fit1$residuals)
# plot of residuals VS. quantiles of a normal distribution, normality? 
qqnorm(fit1$residuals)
```

Sometimes variance of y is not constant across the range of x (heteroscedasticity), Little effect on point estimates but variance estimates will be incorrect, This affects confidence intervals and p-values. 

To account for heteroscedasticity we can, we can
      
* use robust standard error
      
* transform the data
      
* Fit a model that does not assume constant variance (GLM)
      
Some reasons for using data transformations
      
* Original data suggest nonlinearity
      
* Equal variance assumption violated
      
* Normality assumption violated
      
  





