---
title: "linear_regression-shrinkage_estimation"
author: "Ruijuan Li"
date: "7/27/2017"
output: 
  html_document: 
    keep_md: yes
---

# linear regression 
```{r}
# install all the required packages 
# install.packages("gee")
# install.packages("multcomp")
library(gee)
library(multcomp)
```

# lab1, association between triglycerides & BMI??? 
https://github.com/leejimmy93/summer_institute/blob/master/module_4_regression_and_ANOVA/2016_SISG_4_2.pdf (slides 1-58)
```{r}
cholesterol = read.table("http://faculty.washington.edu/rhubb/sisg/SISG-Data-cholesterol.txt", header=T)
attach(cholesterol) # make variables global available in R studio 
# compute univariate summary statistics for triglycerides
dim(cholesterol) # 400 

# take a look at the data 
mean(TG)
median(TG)
sd(TG)
summary(TG)
# graphical displays for triglycerides
boxplot(TG)
stem(TG) # ??? similar to histogram, but in a numeric pattern 
?stem
hist(TG)

# creat a binary indicator for BMI > 25 
ibmi <- ifelse(BMI > 25, 1, 0)  # arbitrarily devide BMI into 2 groups 
ibmi

# compute univariate summary statistics for triglycerides for BMI > 25 and BMI <= 25
tapply(TG,ibmi,mean)
tapply(TG,ibmi,median)
tapply(TG,ibmi,sd)
# tapply(vector, index, function)

# plot boxplots for triglycerides separately by BMI > 25 and BMI <= 25
boxplot(TG ~ ibmi) # summary boxplot for the 2 groups with different BMI 
t.test(TG ~ ibmi) # doing t-test to see whether there is association 

# instead of t-test, using continous predictor and outcome values to test for association 
# scatterplot of triglycerides vs BMI
plot(BMI, TG)
# fit linear regression models for the association between triglycerdies and BMI
fit1 = lm(TG ~ BMI)
?glm # more generlized, can use glm for lm 
summary(fit1)
confint(fit1) # confidence interval 
fit1$coefficients 

# compute the predicted value and its 95% confidence interval for the mean value of TG at BMI=23 as well
# as for a new individual w/ BMI=23. 
predict(fit1, newdata = data.frame(BMI = 23), interval = "confidence") # predicted value and its 95% confidence interval for the mean value of TG at BMI=23
predict(fit1, newdata = data.frame(BMI = 23), interval = "prediction") # predicted value and its 95% confidence interval for a new individual w/ BMI=23

# How do these two intervals differ and why?
# the predicted value for these two are the same but the intervals for predicted value for new individuals are much bigger, because it includes the random variability around the mean. 
```

# check residual plot 
http://www.r-tutor.com/elementary-statistics/simple-linear-regression/residual-plot

# check goodness of fit using R squared
http://blog.minitab.com/blog/adventures-in-statistics-2/regression-analysis-how-do-i-interpret-r-squared-and-assess-the-goodness-of-fit

summary(fit) gives R squred value directly 

# shrinkage 

shrinkage estimator: a naive or raw estimate is improved by combining it with other information. The term relates to the notion that the improved estimate is made closer to the value supplied by the 'other information' than the raw estimate. 

shrinkage & bayes estimation 
